---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.height = 2
)
```

# foodwebr

<!-- badges: start -->
<!-- badges: end -->

`foodwebr` makes it easy to visualise the dependency graph of a set of functions (i.e. who calls
who). This can be useful for exploring an unfamiliar codebase, or reminding yourself what you wrote
ten minutes ago

## Installation

You can install foodwebr from GitHub:

``` r
devtools::install_github("lewinfox/foodwebr")
```

## Basic usage

Say we have a bunch of functions in the global environment, some of which call each other:

```{r setup}
library(foodwebr)

f <- function() 1
g <- function() f()
h <- function() { f(); g() }
i <- function() { f(); g(); h() }
j <- function() j()
```

A call to `foodweb()` will calculate a graph of the dependencies.

```{r foodweb-basic}
fw <- foodweb()
```

Printing the object will show the [graphviz](https://graphviz.org/) representation:

```{r foodweb-print}
fw
```

Plotting will draw the graph.

```{r foodweb-plot, message=FALSE}
plot(fw)
```

`foodweb()` looks at its calling environment by default. If you want to look at another environment
you can either pass a function to the `FUN` argument of `foodweb()` or pass an environment to the
`env` argument. If `FUN` is provided then the value of `env` is ignored, and the environment of
`FUN` will be used.

### Filtering

If a specific function is passed to `FUN`, the default behaviour is to remove functions that are not
descendants or antecedents of that function.

```{r foodweb-filter}
# `j()` will not be included
foodweb(FUN = g)

# Force inclusion of unconnected functions by using `filter = FALSE`
foodweb(FUN = g, filter = FALSE)
```

You can use this feature when exploring code in other packages: calling `foodweb()` on a function in
another package will show you how functions in that package relate to each other. I'm using `cowsay`
here as it's small enough that the output is readable.

By default when calling `foodweb()` on a specific function we only see functions that are in the
direct line of descendants or antecendents of the specified function.

```{r foodweb-plot-package}
if (requireNamespace("cowsay", quietly = TRUE)) {
  plot(foodweb(cowsay::say))
}
```

If we want to include _all_ functions in the package, we can pass `filter = FALSE`:

```{r foodweb-plot-package-no-filter}
if (requireNamespace("cowsay", quietly = TRUE)) {
  plot(foodweb(cowsay::say, filter = FALSE))
}
```

### `graphviz` as text

In case you want to do something with the [graphviz](https://graphviz.org/) output (make it
prettier, for example), use `as.text = TRUE`. This returns the graphviz specification as a character
vector.

```{r foodweb-as-text}
foodweb(as.text = TRUE)
```

Calling `as.character()` on a `foodweb` object will have the same effect.


## Using `tidygraph`

The [`tidygraph`](https://tidygraph.data-imaginist.com/) package provides tools for graph analysis.
A `foodweb` object can be converted into a tidy graph object using `tidygraph::as_tbl_graph()` to
allow more sophisticated analysis and visualisation.

```{r foodweb-tidygraph}
if (requireNamespace("tidygraph", quietly = TRUE)) {
  tg <- tidygraph::as_tbl_graph(foodweb())
  tg
}
```
