---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.height = 2
)
```

# foodwebr

<!-- badges: start -->
<!-- badges: end -->

`foodwebr` makes it easy to visualise the dependency graph of a set of functions (i.e. who calls
who). This can be useful for exploring an unfamiliar codebase, or reminding yourself what you wrote
ten minutes ago

## Installation

You can install foodwebr from GitHub:

``` r
devtools::install_github("lewinfox/foodwebr")
```

## Basic usage

Say we have a bunch of functions in the global environment, some of which call each other:

```{r setup}
library(foodwebr)

f <- function() 1
g <- function() f()
h <- function() { f(); g() }
i <- function() { f(); g(); h() }
j <- function() j()
```

A call to `foodweb()` will calculate a graph of the dependencies, i.e. who calls who.

```{r, foodweb-basic}
fw <- foodweb()
```

Printing the object will show the graphviz representation:

```{r, foodweb-print}
fw
```

Plotting will draw the graph.

```{r foodweb-plot, message=FALSE}
plot(fw)
```

`foodweb()` looks at the global environment by default. If you want to look at another environment
you can either pass a function to the `FUN` argument of `foodweb()` or pass an environment to the
`env` argument. If `FUN` is provided then the value of `env` is ignored, and the environment of
`FUN` will be used.

You can use this to map all the functions in a package. I'm using `cowsay` here as it's small
enough that the output is readable.

```{r, foodweb-plot-package}
if (requireNamespace("cowsay", quietly = TRUE)) {
  plot(foodweb(cowsay::say))
}
```

In case you want to do something with the [graphviz](https://graphviz.org/) output (make it
prettier, for example), use `as.text = TRUE`. This returns the graphviz specification as a character
vector.

```{r foodweb-as-text}
foodweb(as.text = TRUE)
```

## Digging deeper

`foodwebr` also exposes the workhorse functions in case you want to play around with them.

### `foodweb_matrix()`

The starting point is to compute the function matrix. This idea, and much of the implementation,
was taken from [`mvbutils::foodweb()`](https://rdrr.io/cran/mvbutils/man/foodweb.html). The function
matrix is 1 if the function on the y-axis calls the function on the x-axis, and 0  otherwise.
`foodweb_matrix()` looks at functions in the global environment by default, but you can specify
another environment using the `env` argument.

```{r, foodweb-funmat}
funmat <- foodweb_matrix()

funmat
```

Note that self-calls are ignored (`funmat["j", "j"]` is zero even though `j()` calls itself).

### `graphviz_spec_from_matrix()`

`graphviz_spec_from_matrix()` translates the function matrix into a character string containing a
[graphviz](https://graphviz.org/) specification:

```{r, foodweb-graph-spec}
graphviz_spec <- graphviz_spec_from_matrix(funmat)

graphviz_spec
```

### Visualisation

We can visualise the graph specification using `Diagrammer::grViz()`.

```{r foodweb-plot-graph-spec}
DiagrammeR::grViz(graphviz_spec)
```


## Using `tidygraph`

The [`tidygraph`](https://tidygraph.data-imaginist.com/) package provides tools for graph analysis.
A `foodweb` object can be converted into a tidy graph object using `tidygraph::as_tbl_graph()` to
allow more sophisticated analysis and visualisation.

```{r foodweb-tidygraph}
if (requireNamespace("tidygraph", quietly = TRUE)) {
  tg <- tidygraph::as_tbl_graph(foodweb())
  tg
}
```
